The Emergent Simplicity of Galaxy Size

### 2017-09-15 paper/galsize_paper.tex
Replaced figures

In \S\ref{subsec:tests1} we show comparisons between the galaxy size model described in \S\ref{sec:model} and our SDSS sample. We identify the key ingredients that determine the characteristic $\rhalf-$dependence of galaxy clustering in \S\ref{subsec:tests2}. In so doing, we demonstrate the sensitivity of galaxy clustering measurements to the underlying model assumptions, establishing the success of our model as non-trivial.

\subsection{Tests of the Fiducial Power Law Model}
\label{subsec:tests1}

In Figure \ref{fig:scatter_plot} we show the scaling of galaxy size $\rhalf$ with $\mstar.$ 
%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/single_component_model_vs_sdss_one_point.pdf}
\caption{
{\bf One-point data used to fit the fiducial model.}
Scattered points show the $\rhalf-\mstar$ relation for SDSS galaxies as measured in \citet{meert_etal15}. The black curve in each panel shows the $\rhalf-\mstar$ relation implied by the model. This figure confirms the findings in \citet{kravtsov13} that a linear relationship between $\rvir$ and $\rhalf,$ convolved against the nonlinear relationships between $\rvir, \mhalo$ and $\mstar,$ correctly predicts the characteristic curvature in the relation $\langle\rhalf\vert\mstar\rangle$ over a wide range in stellar mass.
}
\label{fig:scatter_plot}
\end{figure}
%-----------------------------------------------------------------------------------------------------

The results shown in \citet{kravtsov13} simplify our task of parameter space exploration: the assumption that composite galaxy size $\rhalf$ scales in linear proportion to halo radius $\rvir$ gives a qualitatively good description of SDSS data over many orders of magnitude in galaxy mass. As first pointed out in \citet{kravtsov13}, this linearization is non-trivial because of the well-known curvature in the scaling relation of $\langle\rhalf\vert\mstar\rangle,$ which is visually apparent in all panels of Fig.~\ref{fig:scatter_plot}.

Comparison of the top panels of Fig.~\ref{fig:scatter_plot} show that this linearization is only approximate; evidently, galaxy size $\rhalf$ of bulge-dominated systems has a steeper scaling relation with $\mstar$ relative to disk-dominated systems. This difference motivates our simple extension of the \citet{kravtsov13} model, in which $\rhalf$ of disk- and bulge-dominated galaxies scales according to separate power laws.

Through MCMC parameter space exploration aided by {\tt emcee} \citep{emcee_hammer}, we find a good description of the observed one-point scaling relations is given by model galaxies with the parameters $\adisk=0.014=7\abulge, \alphadisk=1, \alphabulge=5/4,$ with log-normal scatter of $0.2$ dex motivated by \citet{somerville_etal17}. The black curves in Fig.~\ref{fig:scatter_plot} illustrate the median $\langle\rhalf\vert\mstar\rangle.$

We confirm the visual impression of comparing the top panels of Fig.~\ref{fig:scatter_plot}: we are unable to find a parameter combination in which the composite size of both bulge- and disk-dominated galaxies have linear ($\alpha=1$) scaling with $\rvir.$ We do not quantitatively rule out such a scaling relation, as this would require careful investigation of the systematic error budget of galaxy profile modeling that we consider beyond the scope, though well-motivated, by the present work.

\subsubsection{Dependence of clustering on $\rhalf$}
\label{subsubsec:clustering_tests}

@@

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/size_clustering_ratios_bt_decomposition_model_vs_sdss.pdf}
\caption{
{\bf Two-point clustering predictions of the fiducial model.}
Points with error bars show new SDSS measurements of the $\rhalf-$dependence of projected galaxy clustering, $\wproj,$ compared to predictions by the model tuned to the measurements shown in Fig.~\ref{fig:scatter_plot}. We define a disk or bulge as ``large" or ``small" according to whether it is above or below the median size for its stellar mass. The y-axis shows clustering strength ratios, so that, for example, a y-axis value of $-0.5$ corresponds to small galaxies being $50\%$ more strongly clustered than large galaxies of comparable stellar mass. We show results separately for disk-dominated galaxies ({\em top} panels) and bulge-dominated galaxies ({\em bottom }panels), and different thresholds in total stellar mass in the {\em left} and {\em right} panels. The successful prediction shown here is remarkable because the model was not fit to these data, and because two-point clustering is highly sensitive to the physics that shapes satellite galaxy profiles (see Fig.~\ref{fig:satellites}).
}
\label{fig:clustering_ratio_upshot}
\end{figure}
%-----------------------------------------------------------------------------------------------------

@@

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/random_bt_centrals_vs_satellites.pdf}
\caption{
{\bf Morphologies of centrals and satellites of the same halo mass.} We show the PDF of the bulge-to-total stellar mass ratio $\bt$ of central and satellite galaxies of the same halo mass $\mhalo=\mpeak.$ In the model, the PDF of $\bt$ is determined by the stellar mass $\mstar$ and specific star-formation rate $\ssfr$ statistically determine, with no residual dependence on the cosmic web. Satellite galaxies in the model are ``bulgier" than centrals of the same halo mass because satellites are more quiescent than centrals of the same stellar mass. Thus our baseline model ansatz is that the morphology-density relation is purely derived from the color-density relation. Since composite size $\rhalf$ in the model is given by $\rhalf\equiv\rhalf=(\bt)\rhalfbulge + (1-\bt)\rhalfdisk,$ this ansatz in turn implies that smaller galaxies cluster more strongly relative to larger galaxies of the same stellar mass, as seen in the measurements shown in Fig.~\ref{fig:clustering_ratio_upshot}.
}
\label{fig:bt_censat}
\end{figure}
%-----------------------------------------------------------------------------------------------------

In the {\tt UniverseMachine}, and in any quantitatively successful cosmological model of galaxy formation, central galaxies are preferentially more star-forming relative to satellites of comparable stellar or halo mass. Since bluer galaxies are diskier than redder galaxies, our model centrals will be diskier than satellites. Our galaxy profile model has $\adisk=7\abulge,$ and so satellite galaxies will be smaller than centrals due to the $\bt-\ssfr$ correlation. These simultaneous features alone will naturally result in the negative sign of the clustering ratios seen in Figure~\ref{fig:clustering_ratio_upshot}.

@@

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/alternate_satellite_models_size_clustering_ratios.pdf}
\caption{
{\bf Clustering tightly constrains satellite-specific size evolution.}
Here we compare our fiducial model, in which satellite galaxy size is set by $\rvir$ at the time of infall, to an alternative model analogous to \citet{watson_etal12} in which satellite sizes contract in proportion to $(\mvir/\macc)^{1/3}.$ The large differences between solid and dashed curves in the top panels show that the $\rhalf-$dependence of galaxy clustering ratios is highly sensitive to the post-infall evolution of satellite galaxy profiles. The successful prediction of our fiducial model, in which satellite galaxies neither contract nor puff up after infall, places tight constraints on satellite-specific physical processes, which must be either negligible or conspiratorially produce little-to-no size change after accretion.
}
\label{fig:satellites}
\end{figure}
%-----------------------------------------------------------------------------------------------------

@@

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/size_clustering_ratios_bt_decomposition_spin_size_correlation.pdf}
\includegraphics[width=8cm]{FIGS/central_lensing_spin_size_correlation.pdf}
\caption{
{\bf Clustering and lensing provide no constraining power on the assumption that $\rhalfdisk\propto\halospin.$}
We compare the predictions between our fiducial model in which sizes are purely stochastic, and an alternative model motivated by \citet{mo_mao_white98} in which central galaxy disk size is maximally correlated with host halo spin at fixed stellar mass \citep[implemented via conditional abundance matching, e.g.,][]{hearin_etal13b}. The tiny differences between the solid and dashed curves imply that conventional large-scale structure measurements cannot even in principle provide compelling evidence pertaining to the assumption that $\rhalfdisk\propto\halospin.$
}
\label{fig:halospin_lensing}
\end{figure}
%-----------------------------------------------------------------------------------------------------

>>>>

%In \S\ref{subsec:tests1} we show comparisons between the galaxy size model described in \S\ref{sec:model} and our SDSS sample. We identify the key ingredients that determine the characteristic $\rhalf-$dependence of galaxy clustering in \S\ref{subsec:tests2}. In so doing, we demonstrate the sensitivity of galaxy clustering measurements to the underlying model assumptions, establishing the success of our model as non-trivial.

In Figure \ref{fig:scatter_plot} we show the scaling of galaxy size $\rhalf$ with $\mstar.$

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/single_component_model_vs_sdss_one_point.pdf}
\caption{
{\bf One-point data used to fit the fiducial model.}
Scattered points show the $\rhalf-\mstar$ relation for SDSS galaxies as measured in \citet{meert_etal15}. The black curve in each panel shows the $\rhalf-\mstar$ relation implied by the model. This figure confirms the findings in \citet{kravtsov13} that a linear relationship between $\rvir$ and $\rhalf,$ convolved against the nonlinear relationships between $\rvir, \mhalo$ and $\mstar,$ correctly predicts the characteristic curvature in the relation $\langle\rhalf\vert\mstar\rangle$ over a wide range in stellar mass.
}
\label{fig:scatter_plot}
\end{figure}
%-----------------------------------------------------------------------------------------------------

@@

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/size_clustering_ratios_stripping_model_vs_sdss.pdf}
\caption{
{\bf Two-point clustering predictions of the fiducial model.}
Points with error bars show new SDSS measurements of the $\rhalf-$dependence of projected galaxy clustering, $\wproj,$ compared to predictions by the model tuned to the measurements shown in Fig.~\ref{fig:scatter_plot}. We define a disk or bulge as ``large" or ``small" according to whether it is above or below the median size for its stellar mass. The y-axis shows clustering strength ratios, so that, for example, a y-axis value of $-0.5$ corresponds to small galaxies being $50\%$ more strongly clustered than large galaxies of comparable stellar mass. We show results separately for disk-dominated galaxies ({\em top} panels) and bulge-dominated galaxies ({\em bottom }panels), and different thresholds in total stellar mass in the {\em left} and {\em right} panels. The successful prediction shown here is remarkable because the model was not fit to these data, and because two-point clustering is highly sensitive to the physics that shapes satellite galaxy profiles (see Fig.~\ref{fig:satellites}).
}
\label{fig:clustering_ratio_upshot}
\end{figure}
%-----------------------------------------------------------------------------------------------------

@@

@@

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/scrambled_size_clustering_ratios.pdf}
\caption{
{\bf Clustering tightly constrains satellite-specific size evolution.}
Here we compare our fiducial model, in which satellite galaxy size is set by $\rvir$ at the time of infall, to an alternative model analogous to \citet{watson_etal12} in which satellite sizes contract in proportion to $(\mvir/\macc)^{1/3}.$ The large differences between solid and dashed curves in the top panels show that the $\rhalf-$dependence of galaxy clustering ratios is highly sensitive to the post-infall evolution of satellite galaxy profiles. The successful prediction of our fiducial model, in which satellite galaxies neither contract nor puff up after infall, places tight constraints on satellite-specific physical processes, which must be either negligible or conspiratorially produce little-to-no size change after accretion.
}
\label{fig:satellites}
\end{figure}
%-----------------------------------------------------------------------------------------------------

@@
